---
title: "Analysis of noveltyVR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Aims
This document checks whether the data is saved correctly and will be useable. For this, I did the task myself.

1. Can all analyses be done that are included in the manuscript?
2. Is the counterbalancing (e.g. selection of words) the for both tasks?


# Libraries used
```{r}
library(plyr)
library(MPTinR)
library(matrixTests)
library(reshape2)
library(knitr)
library(chron)
library(ggplot2)
library(MRColour)
library(cowplot)
library(BayesFactor)
theme_set(theme_grey())
```

Important things:
Calculate time difference between encoding and recognition

# Condition
```{r}
# Variables
nBlock         <- 72
listLength     <- 144
conditions     <- 0:11
t1             <- 'alphabetical'
t2             <- 'living'
conditionTable <- data.frame(condition = conditions,
                             list1 = c(1, 1, 2, 2, 3, 3, 1, 1, 2, 2, 3, 3),
                             list2 = c(2, 3, 1, 3, 1, 2, 2, 3, 1, 3, 1, 2),
                             list3 = c(3, 2, 3, 1, 2, 1, 3, 2, 3, 1, 2, 1),
                             aTask = c(rep(t1, 6), rep(t2, 6)),
                             bTask = c(rep(t2, 6), rep(t1, 6)))

conditionTable$aTask <- as.character(conditionTable$aTask)
conditionTable$bTask <- as.character(conditionTable$bTask)

# Note that the col list1, list2 and list3 do not refer to the actual word lists wordList1.txt etc but only 
# to the 1st, 2nd and 3rd list. They col contains which actuall wordList is assign to that spot. This might be confusing as in the recallSheet list1 etc refer to the wordLists
```


## Loading data
### Encoding task
```{r}
folderLocation  <- 'U:/Projects/noveltyVR/data/ignore_rawData/judgement/'
allFiles        <- paste0(folderLocation, list.files(folderLocation))

encodData <- do.call(rbind, lapply(allFiles, function(x) read.table(x, stringsAsFactors = FALSE, header = TRUE)))
```

Is there a reaction time and accuracy difference?
```{r}
# Invert accuracy for 1, 2, 73, 5
bool     <- encodData$subNum == 1 | encodData$subNum == 2 | encodData$subNum == 73 | encodData$subNum == 5
tempData <- encodData[bool, 'acc']
acc      <- rep(NA, length(tempData))
acc[tempData == 0]     <- 1
acc[tempData == 1]     <- 0
encodData[bool, 'acc'] <- acc

# Aggregate
encodData_agg <- ddply(encodData,
                       c('subNum', 'group', 'counterBalancing', 'task'),
                       summarise, 
                       rt = mean(RT, na.rm = TRUE), 
                       acc = mean(acc, na.rm = TRUE))

# Plot
rtPlot <- ggplot(encodData_agg, aes(x = task, y = rt, colour = task)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'RT by task', y = 'RT in msec', x = 'Task') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')

accPlot <- ggplot(encodData_agg, aes(x = task, y = acc, colour = task)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Accuracy by task', y = 'Accuracy', x = 'Task') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')

plot_grid(rtPlot, accPlot)
```

Note that for the first participant, the responses were not saved correctly and have to be inverted. 

```{r}
rtTest <- ttestBF(encodData_agg$rt[encodData_agg$task == 'living'], 
                  encodData_agg$rt[encodData_agg$task == 'alphabetical'], 
                  paired = TRUE)

acTest <- ttestBF(encodData_agg$acc[encodData_agg$task == 'living'],
                  encodData_agg$acc[encodData_agg$task == 'alphabetical'], 
                  paired = TRUE)

rtTest <- round(as.numeric(as.vector(rtTest)))
acTest <- round(1/as.numeric(as.vector(acTest)), 2)
```

There is strong evidence that RT are faster in the living condition (BF10 = `r rtTest`), however there is also evidence that accuracy does not differ between the condition (BF01 = `r acTest`). 

### Recognition task
```{r}
# Look for all files in that folder and load them.
folderLocation  <- 'U:/Projects/noveltyVR/data/ignore_rawData/recognition/'
allFiles        <- paste0(folderLocation, list.files(folderLocation))

recogData <- do.call(rbind, lapply(allFiles, function(x) read.csv(x, stringsAsFactors = FALSE)))
```



```{r}
# Prepare the data for analysis.
# Correct subNum
recogData$subNum[recogData$subNum == 3] <- 73

# Only the test_part recogTask is relevant. 
recogData <- subset(recogData, test_part == 'recogTask')

# Recode variables (1 means studied)
studied  <- rep(1, dim(recogData)[1])
studied[recogData$task == 'unstudied'] <- 0
recogData$studied <- factor(studied, levels = c(0, 1), c('unstudied', 'studied'))

# Response r= 82, n = 78, f = 70
response <- rep(NA, dim(recogData)[1])
response[recogData$key_press == 82] <- 'r'
response[recogData$key_press == 78] <- 'n'
response[recogData$key_press == 70] <- 'f'
recogData$response <- response

# Calculate accuracy
acc  <- rep(NA, dim(recogData)[1])
acc[studied == 1 & (response == 'r' | response == 'f')] <- 1
acc[studied == 0 & (response == 'n')] <- 1
acc[studied == 1 & (response == 'n')] <- 0
acc[studied == 0 & (response == 'r' | response == 'f')] <- 0
recogData$acc <- acc

# Split data into two each encoding task
recogData_alpha  <- subset(recogData, task == 'alphabetical' | task == 'unstudied')
recogData_living <- subset(recogData, task == 'living' | task == 'unstudied')
```


Are there RT difference between RFN?
```{r}
recogData$rt <- as.numeric(recogData$rt)
recogData_rt <- ddply(recogData, c('subNum', 'response'), summarise, RT = mean(rt))

ggplot(recogData_rt, aes(x = response, y = RT, colour = response)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'RT by recognition response', y = 'RT in msec', x = 'Recognition response') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')


recogData_rt$response <- as.factor(recogData_rt$response)
recogData_rt$subNum <- as.factor(recogData_rt$subNum)

rfnTest <- anovaBF(RT ~ response + subNum, data = recogData_rt, whichRandom = "subNum", progress = FALSE)
rfnTest <-round(as.numeric(as.vector(rfnTest)))
```

This compairson showed a main effect of recognition type (BF10 = `r rfnTest`).

Run MTP analysis
```{r}
recogData_alpha_agg <- ddply(recogData_alpha, c('subNum', 'group', 'condition'),
                             summarise, 
                             studiedRemebered = sum(studied == 'studied'   & response == 'r'),
                             studiedFamiliar  = sum(studied == 'studied'   & response == 'f'),
                             studiedNew       = sum(studied == 'studied'   & response == 'n'),
                             newRemebered     = sum(studied == 'unstudied' & response == 'r'),
                             newFamiliar      = sum(studied == 'unstudied' & response == 'f'),
                             newNew           = sum(studied == 'unstudied' & response == 'n'))

fit_alpha            <- fit.mpt(recogData_alpha_agg[, 4:9], "MPT_RKN.model", n.optim = 50)
fit_alpha_individual <- fit_alpha$parameters$individual
fit_alpha_individual <- as.data.frame(t(fit_alpha_individual[,1,]))

recogData_living_agg <- ddply(recogData_living, c('subNum', 'group', 'condition'),
                              summarise,
                              studiedRemebered = sum(studied == 'studied'   & response == 'r'),
                              studiedFamiliar  = sum(studied == 'studied'   & response == 'f'),
                              studiedNew       = sum(studied == 'studied'   & response == 'n'),
                              newRemebered     = sum(studied == 'unstudied' & response == 'r'),
                              newFamiliar      = sum(studied == 'unstudied' & response == 'f'),
                              newNew           = sum(studied == 'unstudied' & response == 'n'))

fit_living            <- fit.mpt(recogData_living_agg[, 4:9], "MPT_RKN.model", n.optim = 50)
fit_living_individual <- fit_living$parameters$individual
fit_living_individual <- as.data.frame(t(fit_living_individual[,1,]))
```

Combining everything into one data frame

```{r}
participants <- unique(recogData$subNum)

row.names(fit_alpha_individual)  <- NULL
row.names(fit_living_individual) <- NULL

mainData_wide            <- rbind(fit_alpha_individual, fit_living_individual)
mainData_wide$subNum     <- rep(recogData_living_agg$subNum, 2)
mainData_wide$group      <- rep(recogData_living_agg$group, 2)
mainData_wide$encodTask  <- c(rep('alphabetical', length(participants)), rep('living', length(participants)))

# Reshape from wide to long
mainData <- melt(mainData_wide, id.vars = c("subNum", "group", "encodTask"))

# Exclude gr and gf
names(mainData)[4] <- 'parameter'
mainData           <- subset(mainData, parameter != 'gf' & parameter != 'gr')


# Make group factor
mainData$group <- factor(mainData$group, level = 1:2, labels = c('Novelty', 'Control'))

# Arcsine transformation
mainData$trans_value <- asin((mainData$value*2)-1)

# Level of processing effect
hypo0 <- ddply(mainData, c('subNum', 'encodTask'), summarise, trans_value = mean(trans_value))
ggplot(hypo0, aes(x = encodTask, y = trans_value, colour = encodTask)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Level of processing effect', y = 'Transformed probability estimate', x = 'Group') +
  scale_color_mrc(palette = 'primary')

# Hypohesis 1.2
hypo1.2 <- ddply(mainData, c('subNum', 'group'), summarise, trans_value = mean(trans_value))
ggplot(hypo1.2, aes(x = group, y = trans_value, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Hypothesis 1.2', y = 'Transformed probability estimate', x = 'Group') +
  scale_color_mrc(palette = 'primary')

# Hypothesis 2.2
hypo2.2 <- ddply(mainData, c('subNum', 'group', 'encodTask'), summarise, trans_value = mean(trans_value))
ggplot(hypo2.2, aes(x = encodTask, y = trans_value, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Hypothesis 2.2', y = 'Transformed probability estimate', x = 'Encoding task') +
  scale_color_mrc(palette = 'primary')

# Hypothesis 3
hypo3 <- ddply(mainData, c('subNum', 'group', 'parameter'), summarise, trans_value = mean(trans_value))
ggplot(hypo3, aes(x = parameter, y = trans_value, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Hypothesis 3', y = 'Transformed probability estimate', x = 'Encoding task') +
  scale_color_mrc(palette = 'primary')

# Hypothesis 4
hypo4 <- mainData
ggplot(hypo4, aes(x = encodTask, y = trans_value, colour = group)) + 
  facet_grid(.~parameter) +
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Hypothesis 4', y = 'Transformed probability estimate', x = 'Encoding task') +
  scale_color_mrc(palette = 'primary')
```

#### Calculate time in hours between starts of encoding and recognition tasks
```{r}
# Format: "2020-01-27 13:45:57"
# Convert data and time for encoding task
encod_startTime      <- strptime(x = paste(encodData$date, encodData$time), format = "%Y%m%d %H%M%S")
encodData$startTime  <- as.character(encod_startTime)

# Get times sorted for participants
encodData_times <- ddply(encodData, 
                         c('subNum', 'group', 'counterBalancing'), 
                         summarise, encodTime = startTime[1])

# Combine recog to 1 string
recog_startTime <- paste0(recogData$year, 
                          '-', 
                          recogData$month + 1, 
                          '-',
                          recogData$day, 
                          ' ',
                          recogData$hour,
                          ':',
                          recogData$minute)
recogData$startTime <- recog_startTime

# Get times sorted for participants
recogData_times <- ddply(recogData, 
                         c('subNum', 'group', 'condition'), 
                         summarise, recogTime = startTime[1])

# Combine to one data frame
times           <- recogData_times
times$encodTime <- encodData_times$encodTime[encodData_times$subNum %in% recogData_times$subNum]
times$diff      <- as.numeric(difftime(times$recogTime, times$encodTime, units = 'hours'))
kable(times)
```

### Recall data
```{r}
recallSheet <- read.csv('U:/Projects/noveltyVR/data/ignore_rawData/recallData.csv')

# Assign correct conditions to list 1, 2 and 3
recallData <- data.frame(subNum    = rep(recallSheet$subNum, each = 2),
                         group     = rep(recallSheet$group, each = 2),
                         condition = rep(recallSheet$condition, each = 2),
                         encodTask = rep(c('alphabetical', 'living'), dim(recallSheet)[1]),
                         recalled  = rep(NA_integer_,  dim(recallSheet)[1]*2))

# Loop through recallSheet
for(i in 1:dim(recallSheet)[1]){
  # Parse condition from conditioonTable and save in tempCond
  tempCond <- conditionTable[which(recallSheet$condition[i] == conditionTable$condition), ]
  
  # Values from 1st and 2nd list
  val1 <- recallSheet[i, paste0('list', tempCond$list1)]
  val2 <- recallSheet[i, paste0('list', tempCond$list2)]
  
  # Assign to recallData according to the aTask
  if(tempCond$aTask == 'alphabetical'){
    recallData[recallData$subNum == recallSheet$subNum[i] & recallData$encodTask == 'alphabetical', 5] <- val1
    recallData[recallData$subNum == recallSheet$subNum[i] & recallData$encodTask == 'living', 5]       <- val2
  } else {
    recallData[recallData$subNum == recallSheet$subNum[i] & recallData$encodTask == 'alphabetical', 5] <- val2
    recallData[recallData$subNum == recallSheet$subNum[i] & recallData$encodTask == 'living', 5]       <- val1
  }
}

# Make factor
recallData$group <- factor(recallData$group, level = 1:2, labels = c('Novelty', 'Control'))

# Hypothesis 1.1
hypo1.1 <- ddply(recallData, c('subNum', 'group'), summarise, recalled = sum(recalled))

hypo2.1 <- recallData
ggplot(hypo2.1, aes(x = encodTask, y = recalled, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Hypothesis 2.1', y = 'Number of words recalled', x = 'Encoding task') +
  scale_color_mrc(palette = 'primary')
```

### Hypothesis summary
```{r}
# Create dataframe
results <- data.frame(Hypothesis = c('1.1', '1.2', '2.1', '2.2', '3', '4'),
                       BF10 = rep(NA, 6),
                       BF01 = rep(NA, 6))
# Hypothesis 1.1
test1.1 <- ttestBF(x = hypo1.1$recalled[hypo1.1$group == 'Novelty'],
                   y = hypo1.1$recalled[hypo1.1$group == 'Control'],
                   nullInterval = c(-Inf, 0))

results[1, 2] <- as.numeric(as.vector(test1.1[2]))
results[1, 3] <- 1/as.numeric(as.vector(test1.1[2]))

# Hypothesis 1.2
test1.2 <- ttestBF(x = hypo1.2$trans_value[hypo1.2$group == 'Novelty'],
                   y = hypo1.2$trans_value[hypo1.2$group == 'Control'],
                   nullInterval = c(-Inf, 0))

results[2, 2] <- as.numeric(as.vector(test1.2[2]))
results[2, 3] <- 1/as.numeric(as.vector(test1.2[2]))

# Hypothesis 2.1
diff2.1.1 <- hypo2.1[hypo2.1$group == 'Control' & hypo2.1$encodTask == 'living', 'recalled'] - 
             hypo2.1[hypo2.1$group == 'Control' & hypo2.1$encodTask == 'alphabetical', 'recalled']

diff2.1.2 <- hypo2.1[hypo2.1$group == 'Novelty' & hypo2.1$encodTask == 'living', 'recalled'] - 
             hypo2.1[hypo2.1$group == 'Novelty' & hypo2.1$encodTask == 'alphabetical', 'recalled']
test2.1   <- ttestBF(x = diff2.1.1,
                     y = diff2.1.2,
                     nullInterval = c(-Inf, 0))

# Plot differnce
df_diff2.1 <- data.frame(group = c(rep('Control', length(diff2.1.1)), rep('Novelty', length(diff2.1.2))),
                         value = c(diff2.1.1, diff2.1.2))

results[3, 2] <- as.numeric(as.vector(test2.1[2]))
results[3, 3] <- 1/as.numeric(as.vector(test2.1[2]))

# Hypothesis 2.2
diff2.2.1 <- hypo2.2[hypo2.2$group == 'Control' & hypo2.2$encodTask == 'living', 'trans_value'] - 
             hypo2.2[hypo2.2$group == 'Control' & hypo2.2$encodTask == 'alphabetical', 'trans_value']

diff2.2.2 <- hypo2.2[hypo2.2$group == 'Novelty' & hypo2.2$encodTask == 'living', 'trans_value'] - 
             hypo2.2[hypo2.2$group == 'Novelty' & hypo2.2$encodTask == 'alphabetical', 'trans_value']
test2.2   <- ttestBF(x = diff2.2.1,
                     y = diff2.2.2,
                     nullInterval = c(-Inf, 0))

results[4, 2] <- as.numeric(as.vector(test2.2[2]))
results[4, 3] <- 1/as.numeric(as.vector(test2.2[2]))

# Plot differnce
df_diff2.2 <- data.frame(group = c(rep('Control', length(diff2.2.1)), rep('Novelty', length(diff2.2.2))),
                         value = c(diff2.2.1, diff2.2.2))


# Hypothesis 3
diff3.1     <- hypo3[hypo3$group == 'Control' & hypo3$parameter == 'r', 'trans_value'] - 
               hypo3[hypo3$group == 'Control' & hypo3$parameter == 'f', 'trans_value']
diff3.2     <- hypo3[hypo3$group == 'Novelty' & hypo3$parameter == 'r', 'trans_value'] - 
               hypo3[hypo3$group == 'Novelty' & hypo3$parameter == 'f', 'trans_value']
test3     <- ttestBF(x = diff3.1,
                     y = diff3.2)

results[5, 2] <- as.numeric(as.vector(test3))
results[5, 3] <- 1/as.numeric(as.vector(test3))


# Plot differnce
df_diff3 <- data.frame(group = c(rep('Control', length(diff3.1)), rep('Novelty', length(diff3.2))),
                         value = c(diff3.1, diff3.2))

# Hypothesis 4
row.names(mainData) <- as.character(1:dim(mainData)[1])
# (E1M1 – E2M1)
diff4.1 <- mainData[mainData$group == 'Control' & mainData$encodTask == 'living' & mainData$parameter == 'f', 'trans_value'] - 
           mainData[mainData$group == 'Control' & mainData$encodTask == 'alphabetical' & mainData$parameter == 'f', 'trans_value']
# (E1M2 – E2M2)
diff4.2 <- mainData[mainData$group == 'Control' & mainData$encodTask == 'living' & mainData$parameter == 'r', 'trans_value'] - 
           mainData[mainData$group == 'Control' & mainData$encodTask == 'alphabetical' & mainData$parameter == 'r', 'trans_value']

# Difference of differences (E1M1 – E2M1) – (E1M2 – E2M2)
diff4.3     <- diff4.1 - diff4.2

# Group 2
# (E1M1 – E2M1)
diff4.4 <- mainData[mainData$group == 'Novelty' & mainData$encodTask =='living' & mainData$parameter == 'f', 'trans_value'] - 
           mainData[mainData$group == 'Novelty' & mainData$encodTask == 'alphabetical' & mainData$parameter == 'f', 'trans_value']
# (E1M2 – E2M2)
diff4.5 <- mainData[mainData$group == 'Novelty' & mainData$encodTask == 'living' & mainData$parameter == 'r', 'trans_value'] - 
           mainData[mainData$group == 'Novelty' & mainData$encodTask == 'alphabetical' & mainData$parameter == 'r', 'trans_value']
# Difference of differences (E1M1 – E2M1) – (E1M2 – E2M2)
diff4.7     <- diff4.4 - diff4.5

# Test
test4     <- ttestBF(x = diff4.3,
                     y = diff4.7)

results[6, 2] <- as.numeric(as.vector(test4))
results[6, 3] <- 1/as.numeric(as.vector(test4))

df_diff4 <- data.frame(group = c(rep('Control', length(diff4.3)), rep('Novelty', length(diff4.7))),
                         value = c(diff4.3, diff4.7))

# Display table
kable(results)

# Level of processing effect
lvlPro <- ttestBF(hypo0$trans_value[hypo0$encodTask == 'living'],
                  hypo0$trans_value[hypo0$encodTask == 'alphabetical'],
                  paired = TRUE)

lvlPro <- round(as.numeric(as.vector(lvlPro)))
```

As a sanity check, we replicated the level of processing effect with better performance for the living condition (BF10 = `r lvlPro`).

```{r, fig.width = 10}
# Plots
# Hypothesis 1.1
pl_hypo1.1 <- ggplot(hypo1.1, aes(x = group, y = recalled, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Hypothesis 1.1', y = 'Words recalled', x = 'Group') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')
  #theme(legend.position = c(1, 0),
   #    legend.justification = c(1, 0),
  #     legend.key.size = unit(0.2, "cm"),
   #    legend.title = element_blank())

# Hypothesis 1.2
pl_hypo1.2 <- ggplot(hypo1.2, aes(x = group, y = trans_value, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Hypothesis 1.2', y = 'Trans. est. probability', x = 'Group') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')

# Hypothesis 2.1
pl_hypo2.1 <- ggplot(df_diff2.1, aes(x = group, y = value, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Differences: Hypothesis 2.1', y = 'Words recalled', x = 'Group') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')

# Hypothesis 2.2
pl_hypo2.2 <- ggplot(df_diff2.2, aes(x = group, y = value, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Differences: Hypothesis 2.2', y = 'Trans. est. probability', x = 'Group') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')

# Hypothesis 3
pl_hypo3 <- ggplot(df_diff3, aes(x = group, y = value, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Differences: Hypothesis 3', y = 'Trans. est. probability', x = 'Group') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')

# Hypothesis 4
pl_hypo4 <- ggplot(df_diff4, aes(x = group, y = value, colour = group)) + 
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Differences: Hypothesis 4', y = 'Trans. est. probability', x = 'Group') +
  scale_color_mrc(palette = 'primary') +
  theme(legend.position = 'none')

plot_grid(pl_hypo1.1, pl_hypo1.2, pl_hypo2.1, pl_hypo2.2, pl_hypo3, pl_hypo4, nrow = 2)
```


# Is the counterbalancing the same across the tasks?
For instance, are the words were learned in the alphabetical task also labelled as learned as such in the encoding task?

```{r}
participants <- unique(recogData$subNum)
check_alpha  <- c()
check_living <- c()
for(i in 1:length(participants)){
  #Alpha
  words_encode <- encodData[encodData$subNum == participants[i] & encodData$task == 'alphabetical', 'word']
  words_recog  <- recogData[recogData$subNum == participants[i] & recogData$task == 'alphabetical', 'word']
  check_alpha[i] <- mean(words_encode %in% words_recog)
  
  # Living
  words_encode <- encodData[encodData$subNum == participants[i] & encodData$task == 'living', 'word']
  words_recog  <- recogData[recogData$subNum == participants[i] & recogData$task == 'living', 'word']
  check_living[i] <- mean(words_encode %in% words_recog)
}

any(!check_alpha == 1)
any(!check_living == 1)
```

As long both check return false, there is no problem with the scoring with the recogntion data. 


### Questionnaire
```{r}
folderLocation   <- 'U:/Projects/noveltyVR/data/ignore_rawData/questionnaire/'
allFiles         <- paste0(folderLocation, list.files(folderLocation))
questData        <- do.call(rbind, lapply(allFiles, function(x) read.csv(x, stringsAsFactors = FALSE)))
names(questData) <- c('num', 
                      'interNum', 
                      'itemName', 
                      'shortcut', 
                      'loading', 
                      'question', 
                      'leftAnchor', 
                      'middleAnchor', 
                      'rightAnchor', 
                      'copyright',
                      'subNum',
                      'date',
                      'time',
                      'rating',
                      'rt')

# Add group to questionnaire
questData$group <- NA
participants    <- unique(questData$subNum)
for(i in 1:length(participants)){
  questData[questData$subNum == participants[i], 'group'] <- encodData_agg$group[encodData_agg$subNum == participants[i]][1]
}

# Make factor
questData$group <- factor(questData$group, level = 1:2, labels = c('Novelty', 'Control'))

# Plot
ggplot(questData, aes(x = shortcut, y = rating, colour = group)) +
  geom_boxplot(outlier.shape = NA) + #avoid plotting outliers twice
  geom_point(position = position_jitterdodge()) +
  labs(title = 'Questionnaire', y = 'Rating', x = '') +
  scale_color_mrc(palette = 'primary') +
  coord_flip()
```